Estimating the genome-wide contribution of selection to temporal allele frequency change

### 2019-10-08 manuscript/manuscript.tex
incorporated some changes

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2} A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2} A). We find these convergent covariances are
relatively weak, and decay very quickly from an initial value of about 0.1
(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
(95\% CIs [0.0087, 0.015]) within 20 generations. This \gc{suggests
  that while a reasonable fraction of the initial response is shared over the
  replicates while the rapid decay} is consistent with
the primary finding of the original \textcite{Barghi2019-qy} study: that
alternative loci contribute to longer term adaptation in \gc{the} different
replicates.

@@

  \caption{{\bf A}: The convergence correlation, averaged across \gc{\textcite{Barghi2019-qy}} replicate
    pairs, through time. Each line represents the convergence correlation
    $\mathrm{cor}(\Delta p_{s}, \Delta p_{s})$ from a starting reference
    generation $s$ to a later time $t$, which varies along the x-axis; each
    line corresponds to a row of the temporal convergence correlation matrix
    depicted to the right.  \gc{We note that convergent correlation
      for the last timepoint is an outlier; we are unsure as to the cause of this, e.g. it does not appear to be driven by a single pair
of replicates.} {\bf B}: The convergence covariance between
    individual pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf
    C}:  The convergence covariance between individual pairs of replicates in
    \parencite{Castro2019-uk} data, for the two selection lines  (LS1 and LS2)
    and the control (Ctrl); gray CIs are those using the complete dataset, blue
    CIs exclude chromosomes 5 and 10 which harbor the two regions
    \textcite{Castro2019-uk} found to have signals of parallel selection between
    LS1 and LS2.}

@@

\gc{Next,} we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response to
selection of about 5 standard deviations over the course of twenty generations
\parencite{Marchini2014-de,Castro2019-uk}. This study includes two independent
selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control
line (Ctrl).  Consequently, this selection experiment offers a useful control
to test our between-replicate covariances: we expect to see positive
between-replicate covariance in the comparison between the two Longshanks
selection lines, but not between the two pairwise comparisons between the
control line and the two Longshanks lines. We find that this is the case (gray
confidence intervals in Figure \ref{fig:figure-2} C), with the two Longshanks
comparisons to the control line not being significantly different from zero,
while the comparison between the two Longshanks lines is statistically
significantly different from zero (CIs [0.013, 0.037]).

\gc{One} finding in the Longshanks study were two major-effect loci that showed
parallel frequency shifts between the two selection lines: a region harboring
the gene Nkx3-2 known to be involved in limb development, and another region
harboring six other candidate genes. We were curious to what extent our
genome-wide covariances were being driven by these two outlier large-effect
loci, so we excluded them from the analysis. Since we do not know the extent to
which linkage disequilibrium around these large-effect loci affects neighboring
loci, we took the conservative precaution of excluding the entire chromosomes
these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
covariances. We find excluding these large effect loci has little impact on the
confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
indicating that these across-replicate covariances are indeed driven by a large
number of loci. This is consistent with a signal of selection on a
  polygenic trait driving genome-wide change, although we note that
  large-effect loci can contribute to the indirect change at unlinked loci
\parencite{Robertson1961-ho,Santiago1995-hx}.

>>>>

We've calculated the convergent correlation for all rows of the replicate
covariance matrices. Like temporal covariances, we visualize these through time
(Figure \ref{fig:figure-2} A), with each line representing the convergent
correlation from a particular reference generation $s$ as it varies with $t$
(shown on the x-axis). In other words, each of the colored lines corresponds to
the like-colored row of the convergence correlation matrix (upper left in
Figure \ref{fig:figure-2} A). We find these convergent covariances are
relatively weak, and decay very quickly from an initial value of about 0.1
(95\% block bootstrap confidence intervals $[0.094, 0.11]$) to around 0.01
(95\% CIs [0.0087, 0.015]) within 20 generations. This \vb{suggests that while
a reasonable fraction of the initial response is shared over the replicates,
this is followed by a rapid decay, a result} consistent with the primary
finding of the original \textcite{Barghi2019-qy} study: that alternative loci
contribute to longer term adaptation \vb{across} the different replicates.

@@

  \caption{{\bf A}: The convergence correlation, averaged across
    \textcite{Barghi2019-qy} replicate pairs, through time. Each line
    represents the convergence correlation $\mathrm{cor}(\Delta p_{s}, \Delta
    p_{s})$ from a starting reference generation $s$ to a later time $t$, which
    varies along the x-axis; each line corresponds to a row of the temporal
    convergence correlation matrix depicted to the right.  We note that
    convergent correlation for the last timepoint is an outlier; we are unsure
    as to the cause of this, e.g. it does not appear to be driven by a single
    pair of replicates. {\bf B}: The convergence covariance between individual
    pairs of replicates in the \textcite{Kelly2019-dc} data. {\bf C}:  The
    convergence covariance between individual pairs of replicates in
    \parencite{Castro2019-uk} data, for the two selection lines  (LS1 and LS2)
    and the control (Ctrl); gray CIs are those using the complete dataset, blue
  CIs exclude chromosomes 5 and 10 which harbor the two regions
\textcite{Castro2019-uk} found to have signals of parallel selection between
LS1 and LS2.}

@@

Next, we reanalyzed the Longshanks selection experiment, which selected for
longer tibiae length relative to body size in mice, leading to a response to
selection of about 5 standard deviations over the course of twenty generations
\parencite{Marchini2014-de,Castro2019-uk}. This study includes two independent
selection lines, Longshanks 1 and 2 (LS1 and LS2), and an unselected control
line (Ctrl).  Consequently, this selection experiment offers a useful control
to test our between-replicate covariances: we expect to see positive
between-replicate covariance in the comparison between the two Longshanks
selection lines, but not between the two pairwise comparisons between the
control line and the two Longshanks lines. We find that this is the case (gray
confidence intervals in Figure \ref{fig:figure-2} C), with the two Longshanks
comparisons to the control line not being significantly different from zero,
while the comparison between the two Longshanks lines is statistically
significantly different from zero (CIs [0.013, 0.037]).

One finding in the Longshanks study \vb{was that two} major-effect loci showed
parallel frequency shifts between the two selection lines: a region harboring
the gene Nkx3-2 known to be involved in limb development, and another region
harboring six other candidate genes. We were curious to what extent our
genome-wide covariances were being driven by these two outlier large-effect
loci, so we excluded them from the analysis. Since we do not know the extent to
which linkage disequilibrium around these large-effect loci affects neighboring
loci, we took the conservative precaution of excluding the entire chromosomes
these loci reside on (chromosomes 5 and 10), and re-calculating the temporal
covariances. We find excluding these large effect loci has little impact on the
confidence intervals (blue confidence intervals in Figure \ref{fig:figure-2}),
indicating that these across-replicate covariances are indeed driven by a large
number of loci. This is consistent with a signal of selection on a polygenic
trait driving genome-wide change, although we note that large-effect loci can
contribute to the indirect change at unlinked loci
\parencite{Robertson1961-ho,Santiago1995-hx}.
